On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme

### 2020-09-03 Revisiones/PNA_MMP_Responses to the reviewers.tex

\maketitle
%\tableofcontents

We  are grateful to the reviewers for taking their time  to review our work. Their comments have helped us to improve the paper. Changes to the original manuscript are given in  color (\textcolor{red}{red} for corrections, \textcolor{blue}{blue} for modified texts)  in the revised version.  A detailed response to both reviewers is given below.

\section*{Response to Reviewer \#1}
{\it
The authors have presented a promissing approach for MPM to surpass the shock wave propagation in the traditional MPM. The paper was very well written and the results suggested that the proposed approach works.
}
\\
\\
The reviewer's positive commentaries are greatly appreciated.
\\
\textit{However, most of the applications presented in this manuscript are relatively simple and does not really demonstrate the capability of the proposed approach. I suggest the authors to extend their manuscript to predict more challenging problems and compare their results to the original MPM solutions.}\\

Authors agree with the proposal of the reviewer. In the first version, aiming to make a readable manuscript, several examples were not shown. In this revised version, in order to clarify several questions of both reviewers, we have introduced an additional case of study which consists in the simulation of a side impact in an elastic square. This test has been conveniently adapted from the work of Hammerquist \& Nairn \cite{HAMMERQUIST2017724}, where fast dynamics were also studied.

@@

\item \textit{The development is very confusing and would make more sense were it to be tied to 
the way that GIMP methods are usually stated, i.e. that the standard FE shape functions are replaced by weighting functions comprised of the conjugation of a standard shape function and a particle "characteristic" function. Line 155 demonstrates this, i.e. the use of $N_{Ip}$ before explanation.}\\

The development of the MPM approach has been revised considering the proposal of the reviewer.

\item \textit{There is little discussion on some of the pitfalls of using max-ent, e.g. the need for a convex domain, the issue of determining Lagrange Multiplier solutions for points close to boundaries, i.e. the solutions process here at Eqns 37 and 38.}\\

Of course, the employment of the \textit{max-ent} approximants is not free of difficulties. Some requirements such as the need for a convex domain, as well as those derived from its calculation (the determination of the Lagrange Multipliers, specially at the boundaries, or the obtainment of the minimizer of the logarithmic function), make the usage of it a challenging tool.

@@

Additionally, Arroyo \& Ortiz~\cite{Arroyo2006} briefly discussed the existence of non-convex domains, and proposed some solutions to it.  For instance, the possibility of replacing the Euclidean distance $\lVert  x - x_a  \rVert$ in the definition of the shape functions by the length of the shortest path contained within the domain connecting $x$ and $x_a$. Or by decomposing the non-convex domain into convex sub-domains.  On the other hand, this topic has been extensively studied in the context of MLS-based meshfree methods, for instance visibility, diffraction, and constrained path criteria.  These methods are directly applicable to local \textit{max-ent} approximation.

However, the results depicted in this manuscript will strengthen the motivation of its employment in order to mitigate typical MPM problems such as cell-crossing or stress instabilities.

\item \textit{Does the use of max-ent solve the problem of imposing Dirichlet boundary conditions in the MPM?}\\

Unfortunately, local \textit{max-ent} approximation technique only provides an efficient tool to transfer information from particles to nodes and vice versa to the MPM. Hence, the difficulties of imposing Dirichlet boundary conditions (DBC) in the MPM persists since they are a direct consequence of discretising the continuum with material points instead of nodes, where traditionally DBC are imposed. However, any available technique employed to overcome this situation in the standard MPM can be translated to the \textit{max-ent} one thanks to the Kronecker-Delta property inherent to it \cite{Arroyo2006}. See for instance the work of Cortis {\it et al.} (2018) \cite{Cortis_et_al_2017_IJNME}.

@@

These interesting publications will be added to the state of art of the revised manuscript. In the humble opinion of the authors, the XPIC approach consists of the stabilisation of the solution through a novel transference procedure more than a time integration scheme. Nevertheless, the algorithm introduced in this research proposes the stabilisation within the adaptation of the predictor-corrector Newmark algorithm, traditionally employed within FEM schemes, for the MPM framework. Hence, both techniques can be employed simultaneously.

\end{enumerate}

 \textit{The numerical examples are not impressive when the message of the paper is that this is good for fast dynamics problems. Compare to papers such as Ma et al. (2009) IJ Impact Engineering}\\

Contrary to the reviewer's opinion, the provided results are capable to highlight the strong capabilities of the MPM Newmark Predictor-Corrector scheme (NPC). It is important to remark the extension of this research: in order to isolate the strength of the proposed methodology, covering spatial and time discretization, only linear elastic problems are assessed. Although the results presented in Ma {\it et al.} \cite{MA2009272} are impressive, the proposed examples are out of the scope of the present research. In order to reinforce the benefits of the proposed methodology and in analogy with the work of Hammerquist \& Nairn \cite{HAMMERQUIST2017724} (previously cited by the reviewer), a 2D wave propagation problem has been simulated. This new test highlights the capabilities of the NPC to reproduce complex wave patterns under challenging loads such as impacts. While at the simultanesly minimices the presence of numerical noise and preserves the solution.

### 2020-09-03 LME-MPM.tex
Cambios en el manuscrito

\MODIFIED{Unfortunately, local \textit{max-ent} approximation technique only provides an efficient tool to transfer information from particles to nodes and vice versa to the \acrshort{mpm}. Hence, the difficulties of imposing \acrshort{dbc} in the \acrshort{mpm} persists since they are a direct consequence of discretising the continuum with material points instead of nodes, where traditionally \acrshort{dbc} are imposed. However, any available technique employed to overcome this situation in the standard \acrshort{mpm} can be translated to the \textit{max-ent} one thanks to the Kronecker-Delta property inherent to it \cite{Arroyo2006}. See for instance the work of Cortis {\it et al.} (2018) \cite{Cortis2018a}}.

@@

\MODIFIED{The following test is best suited to validate the ability of the discussed
time integration scheme to reproduce 2D wave propagation problem.  It consists of the side impact of a moving elastic square. The square was 50 x 50 mm and modeled with the following elastic parameters :
\begin{itemize} 
\item  Initial density : $20$
\item  Poisson ratio : $0.3$
\item  Elastic modulus : $10^5$
\end{itemize}
}

\MODIFIED{The background set of nodes consists in a cartesian grid with a uniform gap between nodes of 1 mm, and square was discretised with four material points by cell. The courant number is fixed to 0.1 during the whole simulation and $\text{LME}_{4.0}$ shape functions where used. The square initially move horizontally through the grid at $\text{V}_0 = 1.0$ until it hits an obstacle modelled as a set of nodes with  zero velocity prescribed. The collision produce stress waves and cause the square to bounce back.}

@@

\MODIFIED{Although during impact mechanical energy decreases, no real dissipation occurs since after collision the value is again 25 kJ like before collision. Furthermore, kinetic energy has been exactly computed. Hence, this apparently dissipation of mechanical energy is probably due to numerical cancellation of shear terms contribution. }
%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.8\hsize}{!}{
\includegraphics[scale=1]{Figures/Energy_square.png} 
  }
  \caption{Snapshots of velocity magnitudes in an elastic square impacted at a point on the right edge. The top row is for \acrshort{npc} updates and the bottom row is for \acrshort{fe} updates.}
  \label{fig:Energy_impact_square}
\end{figure}
%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{sec:conclusions}
We have proposed in this paper an enhanced methodology which improves \acrfull{mpm} behaviour in fast dynamics problems. Different time and space discretisation techniques have been presented and their performance being compared agains other approaches in two benchmark problems.
First, a novel time integration scheme for
\acrshort{mpm} has been carried out. The \acrshort{npc} has been
  shown to be a highly efficient alternative for demanding fast dynamics problems. In the literature, several approaches of the \acrshort{npc} are proposed in order to compete on accuracy with expensive implicit time integration algorithms. The main achievement of the proposed time discretisation has been observed within the Dyka's bar \ref{sec:dyka-bar}, obtaining lower values of the error than the ones obtained with the \acrfull{fe} scheme.
Also, the procedure employed to design the \acrshort{npc} algorithm opens the
door to revisit a huge variety of time integration schemes developed
originally for \acrshort{fem}, which can be implemented within a
  \acrshort{mpm} framework with the suitable modifications. In
addition to the implementation of enhanced time integration schemes, further
research can be done in the analysis of the good performance of the algorithm within non-linear solutions, such as the coupling between different phases or the non-linear constitutive behaviour of the materials: the employment of this scheme to strengthen the localization capabilities of \acrshort{mpm} for viscoplastic materials is on the scope of this research line. The comparison with traditional time schemes in the non-linear field will improve the formal understanding of the proposed methodology.

>>>>

\MODIFIED{Unfortunately, the employment of \textit{max-ent} approximants is not free of pitfalls. Some requirements such as the need for a convex domain, as well as those derived from its calculation (the determination of the Lagrange Multipliers, specially at the boundaries, or the obtainment of the minimizer of the logarithmic function), make the usage of it a challenging tool. Concerning the convex domain requirement, Arroyo \& Ortiz~\cite{Arroyo2006} briefly discussed the existence of non-convex domains, and proposed some solutions to it. For instance, the possibility of replacing the Euclidean distance $\lVert x - x_a \rVert$ in the definition of the shape functions by the length of the shortest path contained within the domain connecting $x$ and $x_a$. Or by decomposing the non-convex domain into convex sub-domains.}

\MODIFIED{Of course the requirement of convex domain is not exclusive from \textit{max-ent}, it also concerns to the remaining interpolation techniques in the \acrshort{mpm} and other meshfree methods. For instance, this topic has been extensively studied in the context of MLS-based meshfree methods, for instance visibility, diffraction, and constrained path criteria. These methods are directly applicable to local \textit{max-ent} approximation. }

\MODIFIED{Despite these drawbacks, the results depicted in this manuscript will strengthen the motivation of its employment in order to mitigate typical \acrshort{mpm} problems such as cell-crossing or stress instabilities in a wide range of problem.}

@@

\MODIFIED{The following test is best suited to validate the ability of the discussed
time integration scheme to reproduce 2D wave propagation problem.  It consists of the side impact of a moving elastic square. The square was 50 x 50 m and modeled with the following elastic parameters :
\begin{itemize} 
\item  Initial density : $20 kg/m^3$
\item  Poisson ratio : $0.3$
\item  Elastic modulus : $0.1 MPa$
\end{itemize}
}

\MODIFIED{The background set of nodes consists in a cartesian grid with a uniform gap between nodes of 1 mm, and square was discretised with four material points by cell. The courant number is fixed to 0.1 during the whole simulation and $\text{LME}_{4.0}$ shape functions where used. The square initially move horizontally through the grid at $\text{V}_0 = 1.0 m/s$ until it hits an obstacle modelled as a set of nodes with  zero velocity prescribed. The collision produce stress waves and cause the square to bounce back.}

@@

\MODIFIED{Although during impact mechanical energy decreases, no real dissipation occurs since after collision the value is again 25 kJ like before collision. Furthermore, kinetic energy has been exactly computed. Hence, this apparently dissipation of mechanical energy is probably due to numerical cancellation of shear terms contribution. }
%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.8\hsize}{!}{
\includegraphics[scale=1]{Figures/Energy_square.png} 
  }
  \caption{Total energy (kinetic energy plus strain energy) evolution for the square. Solid line represent total energy, dashed line represent strain energy, and dotted line represent kinetic energy.}
  \label{fig:Energy_impact_square}
\end{figure}
%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{sec:conclusions}
We have proposed in this paper an enhanced methodology which improves \acrfull{mpm} behaviour in fast dynamics problems. Different time and space discretisation techniques have been presented and their performance being compared agains other approaches in two benchmark problems.
First, a novel time integration scheme for \acrshort{mpm} has been carried out. The \acrshort{npc} has been shown to be a highly efficient alternative for demanding fast dynamics problems. In the literature, several approaches of the \acrshort{npc} are proposed in order to compete on accuracy with expensive implicit time integration algorithms. The main achievement of the proposed time discretisation has been observed within the Dyka's bar \ref{sec:dyka-bar}, obtaining lower values of the error than the ones obtained with the \acrfull{fe} scheme, \MODIFIED{with a minimal loose of energy as in highlighted in \ref{sec:Velocity-waves-elastic-domain} }.
Also, the procedure employed to design the \acrshort{npc} algorithm opens the
door to revisit a huge variety of time integration schemes developed
originally for \acrshort{fem}, which can be implemented within a \acrshort{mpm} framework with the suitable modifications. In addition to the implementation of enhanced time integration schemes, further research can be done in the analysis of the good performance of the algorithm within non-linear solutions, such as the coupling between different phases or the non-linear constitutive behaviour of the materials: the employment of this scheme to strengthen the localization capabilities of \acrshort{mpm} for viscoplastic materials is on the scope of this research line. The comparison with traditional time schemes in the non-linear field will improve the formal understanding of the proposed methodology.
